---
title: "BST260_Final_Project"
author: "Qingru Xu"
date: "2021/11/30"
output: html_document
---

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(GGally)
```

## Exploratory Data Analysis

```{r}
data <- read_csv("music_genre.csv")
```

```{r}
#Look at data
head(data)
summary(data)
dim(data)
```

```{r}
#Check and remove NAs
data[!complete.cases(data),]
data_clean <- na.omit(data)
dim(data_clean)
```

```{r}
#Check and remove duplicates
data_clean[duplicated(data_clean),]#all NAs, so remove
```

```{r}
#Turn variables into right types
data_clean <- data_clean %>% mutate(
                        key = as.factor(key),
                        mode = as.factor(mode),
                        obtained_date = as.factor(obtained_date),
                        music_genre = as.factor(music_genre),
                        tempo = as.numeric(tempo)
                        )
```
```{r}
#check and remove NAs introduced by tempo error
data_clean[!complete.cases(data_clean),]
data_clean <- na.omit(data_clean)
dim(data_clean)
```

```{r}
#treat date variable
unique(data_clean$obtained_date)

#meaningless for prediction, so remove the column
data_clean <- data_clean %>% select(-c("obtained_date"))

#remove rows with duration_ms variable with -1
data_clean <- data_clean[-which(data_clean$duration_ms == -1),]
```


```{r}
#Look at data
summary(data_clean)
dim(data_clean)
```

```{r}
#visualize artist_name top 10
Artist <- as.data.frame(table(data_clean$artist_name)) %>%
  arrange(desc(Freq))
head(Artist)#first row empty filed should be removed

Artist <- Artist[-1,]
head(Artist)

top10_Artist <- Artist[c(1:10),]
top10_Artist$Var1 <- factor(top10_Artist$Var1, levels = rev(top10_Artist$Var1))
top10_Artist %>%
  ggplot()+
  geom_histogram(aes(x = Var1, y = Freq),stat = "identity")+
  coord_flip()+
  ylab("Number of songs per artist")+
  xlab("Artist name")+
  ggtitle("Top 10 Artists with the most number of songs")+
  theme_bw()
```


```{r}
#instance_id and track_name (unique) are useless for prediction and artist_name too compliated, remove
data_clean <- data_clean %>% select(-c("instance_id","artist_name","track_name"))
```

```{r}
#Categorical Variable visualization
data_clean %>%
  gather(predictor, value, c(key, mode, music_genre)) %>%
  ggplot()+
  geom_bar(aes(x=value), stat = "count")+
  facet_wrap(~ predictor, 
             scales = "free",
             labeller = 
             as_labeller(c("key"="Key", 
                "mode" = "Mode", 
                "music_genre" = "Music Genre")))+
  xlab(NULL)+ylab(NULL)+theme_bw()+
  theme(axis.text.x = element_text(angle = 45,hjust=1))+
  ggtitle("Categorical Variables Distribution")
```

```{r}
#Continuous variables visualization
'%!in%' <- function(x,y)!('%in%'(x,y))
names <- colnames(data_clean)[colnames(data_clean)%!in% c("key", "mode", "music_genre")]

data_clean %>%
  gather(predictor, value, names) %>%
  ggplot()+
  geom_histogram(aes(x=value))+
  facet_wrap(~ predictor, 
             scales = "free",labeller = 
             as_labeller(c("popularity"="Popularity", 
                "acousticness" = "Acousticness", 
                "danceability" = "Danceability",
                "duration_ms" = "Duration (ms)",
                "energy" = "Energy",
                "instrumentalness" = "Instrumentalness",
                "liveness" = "Liveness",
                "loudness" = "Loudness",
                "speechiness" = "Speechiness",
                "tempo" = "Tempo",
                "valence" = "Valence")))+
  xlab(NULL)+ylab(NULL)+theme_bw()+
  ggtitle("Continuous Variables Distribution")
```

```{r}
#Categorical Variable Distribution with Popularity
data_clean %>%
  gather(predictor, value, c(key, mode, music_genre)) %>%
  ggplot()+
  geom_boxplot(aes(x=value, y = popularity))+
  facet_wrap(~ predictor, 
             scales = "free",
             labeller = 
             as_labeller(c("key"="Key", 
                "mode" = "Mode", 
                "music_genre" = "Music Genre")))+
  xlab(NULL)+ylab("Popularity")+theme_bw()+
  theme(axis.text.x = element_text(angle = 45,hjust=1))+
  ggtitle("Categorical Variables Distribution with Popularity")
```

```{r}
#Continuous Variables Distribution with Popularity
data_clean %>%
  gather(predictor, value, names[-1]) %>%
  ggplot()+
  geom_point(aes(x=value, y = popularity), size = 0.001)+
  facet_wrap(~ predictor, 
             scales = "free",
             labeller =
             as_labeller(c(
                "acousticness" = "Acousticness", 
                "danceability" = "Danceability",
                "duration_ms" = "Duration (ms)",
                "energy" = "Energy",
                "instrumentalness" = "Instrumentalness",
                "liveness" = "Liveness",
                "loudness" = "Loudness",
                "speechiness" = "Speechiness",
                "tempo" = "Tempo",
                "valence" = "Valence")))+
  xlab(NULL)+ylab("Popularity")+theme_bw()+
  ggtitle("Continuous Variables Distribution with Popularity")
```

```{r}
#Continuous Variables Distribution with Mode
data_clean %>%
  gather(predictor, value, names) %>%
  ggplot()+
  geom_boxplot(aes(x=mode, y = value))+
  facet_wrap(~ predictor, 
             scales = "free_y",
             labeller =
             as_labeller(c("popularity"="Popularity", 
                "acousticness" = "Acousticness", 
                "danceability" = "Danceability",
                "duration_ms" = "Duration (ms)",
                "energy" = "Energy",
                "instrumentalness" = "Instrumentalness",
                "liveness" = "Liveness",
                "loudness" = "Loudness",
                "speechiness" = "Speechiness",
                "tempo" = "Tempo",
                "valence" = "Valence")))+
  xlab("Mode")+ylab(NULL)+theme_bw()+
  ggtitle("Continuous Variables Distribution with Mode")
```

```{r}
#Continuous Variable Distribution with Music Genre
data_clean %>%
  gather(predictor, value, names) %>%
  ggplot()+
  geom_boxplot(aes(x=music_genre, y = value))+
  facet_wrap(~ predictor, 
             scales = "free_y",
             labeller =
             as_labeller(c("popularity"="Popularity", 
                "acousticness" = "Acousticness", 
                "danceability" = "Danceability",
                "duration_ms" = "Duration (ms)",
                "energy" = "Energy",
                "instrumentalness" = "Instrumentalness",
                "liveness" = "Liveness",
                "loudness" = "Loudness",
                "speechiness" = "Speechiness",
                "tempo" = "Tempo",
                "valence" = "Valence")))+
  xlab("Music Genre")+ylab(NULL)+theme_bw()+
  theme(axis.text.x = element_text(angle = 60,hjust=1))+
  ggtitle("Continuous Variable Distribution with Music Genre")

```

## Machine Learning 

```{r}
#population as outcome (Regression),variable slection, visualization

```

```{r}
#population as outcome (Classification),variable slection, visualization
```

```{r}
#music genre as outcome (Multiclass Classification), variable slection, visualization

```

#### Mode as outcome (binary Classfication), variable slection, visualization

```{r}
library(splitstackshape)
library(caret)
library(e1071)
library(MASS)
library(pROC)
library(rpart)
library(randomForest)
library(knitr)
```

##### Exploratory analysis (EDA)

```{r}
# Frequency of mode
count(data_clean, mode)
```

```{r}
# Contingency table for mode and music genre
tab1 <- xtabs(~ mode + music_genre, data = data_clean)
tab1
```

```{r}
# Box plots for continuous predictors
par(mfrow=c(2,3))
boxplot(popularity ~ mode, data = data_clean)
boxplot(acousticness ~ mode, data = data_clean)
boxplot(danceability ~ mode, data = data_clean)
boxplot(duration_ms ~ mode, data = data_clean)
boxplot(energy ~ mode, data = data_clean)
boxplot(instrumentalness ~ mode, data = data_clean)
boxplot(liveness ~ mode, data = data_clean)
boxplot(loudness ~ mode, data = data_clean)
boxplot(speechiness ~ mode, data = data_clean)
boxplot(tempo ~ mode, data = data_clean)
boxplot(valence ~ mode, data = data_clean)
```

```{r}
# Histogram for continuous predictors
h1 <- data_clean %>% 
  ggplot(aes(popularity, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h2 <- data_clean %>% 
  ggplot(aes(acousticness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h3 <- data_clean %>% 
  ggplot(aes(danceability, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h4 <- data_clean %>% 
  ggplot(aes(duration_ms, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h5 <- data_clean %>% 
  ggplot(aes(energy, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h6 <- data_clean %>% 
  ggplot(aes(instrumentalness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h7 <- data_clean %>% 
  ggplot(aes(liveness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h8 <- data_clean %>% 
  ggplot(aes(loudness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h9 <- data_clean %>% 
  ggplot(aes(instrumentalness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h10 <- data_clean %>% 
  ggplot(aes(liveness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h11 <- data_clean %>% 
  ggplot(aes(loudness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)

grid.arrange(h1,h2,h3,h4,h5,h6,h7,h8,h9,h10,h11, ncol = 4)
```

```{r}
# Stacked bar charts for categorical predictors
p1 <- data_clean %>% ggplot(aes(x=key,fill=factor(mode))) +
  geom_bar(position = "fill") +
  stat_count(geom = "text", 
             aes(label = paste(round((..count..)/sum(..count..)*100), "%")),
             position=position_fill(vjust=0.5), colour="white")
p2 <- data_clean %>% ggplot(aes(x=music_genre,fill=factor(mode))) +
  geom_bar(position = "fill") +
  stat_count(geom = "text", 
             aes(label = paste(round((..count..)/sum(..count..)*100), "%")),
             position=position_fill(vjust=0.5), colour="white")
grid.arrange(p1,p2,ncol=1)
```

According to EDA, four continuous variables (acousticness, danceability, instrumentalness,  speechiness) and two categorical variables (key and music genre) visually display different distributions at two types of modes. So the six variables will be selected as the predictors for the following binary classification. 

##### Prepare the data for binary classification

```{r}
# Set level for mode (Minor=1, Major=0)
levels(data_clean$mode) <- c(0,1)
```

A 70% training, 30% test set split is used. Since there is an imbalance in the outcome groups (25959/40560 = 64% songs with major modes and 14601/40560 = 36% with minor modes), we take this into account during data spliting and use stratified function to split data the same percent of songs from each class (major/minor). 

```{r}
# Create training and test sets
set.seed(1)

x <- stratified(data_clean, "mode", 0.7, keep.rownames = TRUE)
train_set <- x %>% dplyr::select(-rn)
train_index <- as.numeric(x$rn)
test_set <- data_clean[-train_index,]

dim(train_set)
dim(test_set)
```

##### Logistic regression

```{r}
# Fit a logistic regression model
glm_fit <- glm(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data= train_set, family = "binomial")
# Prediction for the test set 
p_hat_logit <- predict(glm_fit, newdata = test_set, type="response")
# Convert the probabilities to predicted response labels 
y_hat_logit <- ifelse(p_hat_logit > 0.5, 1, 0)
cm_logit <- confusionMatrix(data = as.factor(y_hat_logit), reference = as.factor(test_set$mode))
cm_logit
```

##### Naive Bayes

```{r}
# Fit a Naive Bayes model 
nb_fit <- naiveBayes(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_nb <- predict(nb_fit, test_set, type = "raw")[,2]
# Convert the probabilities to predicted response labels 
y_hat_nb <- predict(nb_fit, test_set)
# Confusion matrix
cm_nb <- confusionMatrix(data = as.factor(y_hat_nb), reference = as.factor(test_set$mode))
cm_nb
```

##### kNN

```{r}
# Fit a k-nearest neighbors model
knn_fit <- knn3(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set, k=7)
# Prediction for the test set 
f_hat <- predict(knn_fit, newdata = test_set)[,2]
# Confusion matrix
f_tab <- table(pred=round(f_hat), truth=test_set$mode)
cm_knn <- confusionMatrix(f_tab)
cm_knn
```

##### LDA

```{r}
# Fit an LDA model 
set.seed(1)
lda_fit <- lda(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_lda <- predict(lda_fit, newdata = test_set)$posterior[,2]
# Convert the probabilities to predicted response labels 
y_hat_lda <- ifelse(p_hat_lda > 0.5, 1, 0)
# Confusion matrix
cm_lda <- confusionMatrix(data = as.factor(y_hat_lda), reference = as.factor(test_set$mode))
cm_lda
```

##### QDA

```{r}
# Fit a QDA model
set.seed(1)
qda_fit <- qda(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_qda <- predict(qda_fit, newdata = test_set)$posterior[,2]
# Convert the probabilities to predicted response labels 
y_hat_qda <- ifelse(p_hat_qda > 0.5, 1, 0)
# Confusion matrix
cm_qda <- confusionMatrix(data = as.factor(y_hat_qda), reference = as.factor(test_set$mode))
cm_qda
```

##### Decision trees

```{r}
# Fit a decision tree model 
set.seed(1)
rpart_fit <- rpart(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_rpart <- predict(rpart_fit, newdata = test_set)[,2]
# Convert the probabilities to predicted response labels 
y_hat_rpart <- ifelse(p_hat_rpart > 0.5, 1, 0)
# Confusion matrix
cm_rpart <- confusionMatrix(data = as.factor(y_hat_rpart), reference = as.factor(test_set$mode))
cm_rpart
```

##### Random forest

```{r}
# Fit a random forest model 
set.seed(1)
rf_fit <- randomForest(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_rf <- predict(rf_fit, newdata = test_set, type = "prob")[,2]
# Convert the probabilities to predicted response labels 
y_hat_rf <- ifelse(p_hat_rf > 0.5, 1, 0)
# Confusion matrix
cm_rf <- confusionMatrix(data = as.factor(y_hat_rf), reference = as.factor(test_set$mode))
cm_rf
```

```{r}
# Variable importance table
variable_importance <- importance(rf_fit) 
tmp <- tibble(feature = rownames(variable_importance),
                  Gini = variable_importance[,1]) %>%
                  arrange(desc(Gini))
kable(tmp)
```

```{r}
# Bar plot of variable importance 
tmp %>% filter(Gini > 200) %>%
        ggplot(aes(x=reorder(feature, Gini), y=Gini)) +
        geom_bar(stat='identity') +
        coord_flip() + xlab("Feature") +
        theme(axis.text=element_text(size=8))
```

##### Model comparison

###### Overall accuracy, specificity, and sensitivity

```{r}
cm_logit$byClass[1:2]
```

```{r}
tab2 <- matrix(c(cm_logit$overall[1],cm_logit$byClass[1:2],
                 cm_nb$overall[1],cm_nb$byClass[1:2],
                 cm_knn$overall[1],cm_knn$byClass[1:2],
                 cm_lda$overall[1],cm_lda$byClass[1:2],
                 cm_qda$overall[1],cm_qda$byClass[1:2],
                 cm_rpart$overall[1],cm_rpart$byClass[1:2],
                 cm_rf$overall[1],cm_rf$byClass[1:2]), ncol=3, byrow = TRUE)
colnames(tab2) <- c("Accuracy","Sensitivity","Specificity")
rownames(tab2) <- c("Logistic Regression", "Naive Bayes", "kNN", "LDA", "QDA", "Decision Tree", "Random Forest")
tab2
```

###### ROC curves

```{r}
## Logistic regression
roc_glm <- roc(as.factor(test_set$mode),p_hat_logit)
## Naive Bayes
roc_nb <- roc(as.factor(test_set$mode), p_hat_nb)
## kNN
roc_knn <- roc(as.factor(test_set$mode), f_hat)
## LDA
roc_lda <- roc(as.factor(test_set$mode),p_hat_lda)
## QDA
roc_qda <- roc(as.factor(test_set$mode), p_hat_qda)
## Decision tree
roc_rpart <- roc(as.factor(test_set$mode), p_hat_rpart)
## Random forest
roc_rf <- roc(as.factor(test_set$mode), p_hat_rf)
# Graph with 7 ROC curves for each model
ggroc(list("Logistic regression" = roc_glm, "Naive Bayes" = roc_nb, "kNN" = roc_knn, "LDA" = roc_lda, "QDA" = roc_qda, "Decision tree" = roc_rpart, "Random forest" = roc_rf)) +
  theme(legend.title = element_blank()) +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color = "black", linetype = "dashed") +
  xlab("Sensitivity") +
  ylab("Specificity") 
```

###### AUC values

```{r}
auc(roc_glm)
auc(roc_nb)
auc(roc_knn)
auc(roc_lda)
auc(roc_qda)
auc(roc_rpart)
auc(roc_rf)
```

