---
title: "BST260_Final_Project"
author: "Wenyu Jin"
date: "2021/11/30"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

### Machine Learning I (Binary Classification)
#### Mode (binary variable) as outcome

```{r}
library(splitstackshape)
library(caret)
library(e1071)
library(MASS)
library(pROC)
library(rpart)
library(randomForest)
library(knitr)
library(dplyr)
library(ggplot2)
library(gridExtra)
```

```{r}
data_clean <- readRDS("music_genre_clean.rds")
```

##### Exploratory analysis (EDA)

```{r}
# Frequency of mode
count(data_clean, mode)
```

```{r}
# Contingency table for mode and music genre
tab1 <- xtabs(~ mode + music_genre, data = data_clean)
tab1
```

```{r}
# Box plots for continuous predictors
par(mfrow=c(2,3))
boxplot(popularity ~ mode, data = data_clean)
boxplot(acousticness ~ mode, data = data_clean)
boxplot(danceability ~ mode, data = data_clean)
boxplot(duration_ms ~ mode, data = data_clean)
boxplot(energy ~ mode, data = data_clean)
boxplot(instrumentalness ~ mode, data = data_clean)
boxplot(liveness ~ mode, data = data_clean)
boxplot(loudness ~ mode, data = data_clean)
boxplot(speechiness ~ mode, data = data_clean)
boxplot(tempo ~ mode, data = data_clean)
boxplot(valence ~ mode, data = data_clean)
```

```{r}
# Histogram for continuous predictors
h1 <- data_clean %>% 
  ggplot(aes(popularity, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h2 <- data_clean %>% 
  ggplot(aes(acousticness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h3 <- data_clean %>% 
  ggplot(aes(danceability, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h4 <- data_clean %>% 
  ggplot(aes(duration_ms, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h5 <- data_clean %>% 
  ggplot(aes(energy, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h6 <- data_clean %>% 
  ggplot(aes(instrumentalness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h7 <- data_clean %>% 
  ggplot(aes(liveness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h8 <- data_clean %>% 
  ggplot(aes(loudness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h9 <- data_clean %>% 
  ggplot(aes(instrumentalness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h10 <- data_clean %>% 
  ggplot(aes(liveness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)
h11 <- data_clean %>% 
  ggplot(aes(loudness, ..density..)) +
  geom_histogram(bins=10, color="black") +
  facet_grid(mode~.)

grid.arrange(h1,h2,h3,h4,h5,h6,h7,h8,h9,h10,h11, ncol = 4)
```

```{r}
# Stacked bar charts for categorical predictors
p1 <- data_clean %>% ggplot(aes(x=key,fill=factor(mode))) +
  geom_bar(position = "fill") +
  stat_count(geom = "text", 
             aes(label = paste(round((..count..)/sum(..count..)*100), "%")),
             position=position_fill(vjust=0.5), colour="white")
p2 <- data_clean %>% ggplot(aes(x=music_genre,fill=factor(mode))) +
  geom_bar(position = "fill") +
  stat_count(geom = "text", 
             aes(label = paste(round((..count..)/sum(..count..)*100), "%")),
             position=position_fill(vjust=0.5), colour="white")
grid.arrange(p1,p2,ncol=1)
```

According to EDA, four continuous variables (acousticness, danceability, instrumentalness,  speechiness) and two categorical variables (key and music genre) visually display different distributions at two types of modes. So the six variables will be selected as the predictors for the following binary classification. 

##### Prepare the data for binary classification

```{r}
# Set level for mode (Minor=1, Major=0)
levels(data_clean$mode) <- c(0,1)
```

A 70% training, 30% test set split is used. Since there is an imbalance in the outcome groups (25959/40560 = 64% songs with major modes and 14601/40560 = 36% with minor modes), we take this into account during data spliting and use stratified function to split data the same percent of songs from each class (major/minor). 

```{r}
# Create training and test sets
set.seed(1)

x <- stratified(data_clean, "mode", 0.7, keep.rownames = TRUE)
train_set <- x %>% dplyr::select(-rn)
train_index <- as.numeric(x$rn)
test_set <- data_clean[-train_index,]

dim(train_set)
dim(test_set)
```

##### Logistic regression

```{r}
# Fit a logistic regression model
glm_fit <- glm(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data= train_set, family = "binomial")
# Prediction for the test set 
p_hat_logit <- predict(glm_fit, newdata = test_set, type="response")
# Convert the probabilities to predicted response labels 
y_hat_logit <- ifelse(p_hat_logit > 0.5, 1, 0)
cm_logit <- confusionMatrix(data = as.factor(y_hat_logit), reference = as.factor(test_set$mode))
cm_logit
```

##### Naive Bayes

```{r}
# Fit a Naive Bayes model 
nb_fit <- naiveBayes(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_nb <- predict(nb_fit, test_set, type = "raw")[,2]
# Convert the probabilities to predicted response labels 
y_hat_nb <- predict(nb_fit, test_set)
# Confusion matrix
cm_nb <- confusionMatrix(data = as.factor(y_hat_nb), reference = as.factor(test_set$mode))
cm_nb
```

##### kNN

```{r}
# Fit a k-nearest neighbors model
knn_fit <- knn3(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set, k=7)
# Prediction for the test set 
f_hat <- predict(knn_fit, newdata = test_set)[,2]
# Confusion matrix
f_tab <- table(pred=round(f_hat), truth=test_set$mode)
cm_knn <- confusionMatrix(f_tab)
cm_knn
```

##### LDA

```{r}
# Fit an LDA model 
set.seed(1)
lda_fit <- lda(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_lda <- predict(lda_fit, newdata = test_set)$posterior[,2]
# Convert the probabilities to predicted response labels 
y_hat_lda <- ifelse(p_hat_lda > 0.5, 1, 0)
# Confusion matrix
cm_lda <- confusionMatrix(data = as.factor(y_hat_lda), reference = as.factor(test_set$mode))
cm_lda
```

##### QDA

```{r}
# Fit a QDA model
set.seed(1)
qda_fit <- qda(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_qda <- predict(qda_fit, newdata = test_set)$posterior[,2]
# Convert the probabilities to predicted response labels 
y_hat_qda <- ifelse(p_hat_qda > 0.5, 1, 0)
# Confusion matrix
cm_qda <- confusionMatrix(data = as.factor(y_hat_qda), reference = as.factor(test_set$mode))
cm_qda
```

##### Decision trees

```{r}
# Fit a decision tree model 
set.seed(1)
rpart_fit <- rpart(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_rpart <- predict(rpart_fit, newdata = test_set)[,2]
# Convert the probabilities to predicted response labels 
y_hat_rpart <- ifelse(p_hat_rpart > 0.5, 1, 0)
# Confusion matrix
cm_rpart <- confusionMatrix(data = as.factor(y_hat_rpart), reference = as.factor(test_set$mode))
cm_rpart
```

##### Random forest

```{r}
# Fit a random forest model 
set.seed(1)
rf_fit <- randomForest(mode ~ acousticness + danceability + instrumentalness + speechiness + key + music_genre, data=train_set)
# Prediction for the test set 
p_hat_rf <- predict(rf_fit, newdata = test_set, type = "prob")[,2]
# Convert the probabilities to predicted response labels 
y_hat_rf <- ifelse(p_hat_rf > 0.5, 1, 0)
# Confusion matrix
cm_rf <- confusionMatrix(data = as.factor(y_hat_rf), reference = as.factor(test_set$mode))
cm_rf
```

```{r}
# Variable importance table
variable_importance <- importance(rf_fit) 
tmp <- tibble(feature = rownames(variable_importance),
                  Gini = variable_importance[,1]) %>%
                  arrange(desc(Gini))
kable(tmp)
```

```{r}
# Bar plot of variable importance 
tmp %>% filter(Gini > 200) %>%
        ggplot(aes(x=reorder(feature, Gini), y=Gini)) +
        geom_bar(stat='identity') +
        coord_flip() + xlab("Feature") +
        theme(axis.text=element_text(size=8))
```

Among the six variables, speechiness, acousticness and danceability have Gini indices above 2000 and therefore have relatively higher variable importance compared to the other three variables. So random forest with only speechiness, acousticness and danceability will be fitted to explore whether fewer variables will give similar or better performance. 
```{r}
# Fit another random forest model 
set.seed(1)
rf_fit2 <- randomForest(mode ~ acousticness + danceability + speechiness, data=train_set)
# Prediction for the test set 
p_hat_rf2 <- predict(rf_fit2, newdata = test_set, type = "prob")[,2]
# Convert the probabilities to predicted response labels 
y_hat_rf2 <- ifelse(p_hat_rf2 > 0.5, 1, 0)
# Confusion matrix
cm_rf2 <- confusionMatrix(data = as.factor(y_hat_rf2), reference = as.factor(test_set$mode))
cm_rf2
```

After including the 3 variables with Gini index above 2000, the overall accuracy decreases from 0.717 to 0.646. Thus, we will still stick with the machine learning models containing the 6 variables. 

##### Model comparison

###### Overall accuracy, specificity, and sensitivity

```{r}
cm_logit$byClass[1:2]
```

```{r}
tab2 <- matrix(c(cm_logit$overall[1],cm_logit$byClass[1:2],
                 cm_nb$overall[1],cm_nb$byClass[1:2],
                 cm_knn$overall[1],cm_knn$byClass[1:2],
                 cm_lda$overall[1],cm_lda$byClass[1:2],
                 cm_qda$overall[1],cm_qda$byClass[1:2],
                 cm_rpart$overall[1],cm_rpart$byClass[1:2],
                 cm_rf$overall[1],cm_rf$byClass[1:2]), ncol=3, byrow = TRUE)
colnames(tab2) <- c("Accuracy","Sensitivity","Specificity")
rownames(tab2) <- c("Logistic Regression", "Naive Bayes", "kNN", "LDA", "QDA", "Decision Tree", "Random Forest")
tab2
```

###### ROC curves

```{r}
## Logistic regression
roc_glm <- roc(as.factor(test_set$mode),p_hat_logit)
## Naive Bayes
roc_nb <- roc(as.factor(test_set$mode), p_hat_nb)
## kNN
roc_knn <- roc(as.factor(test_set$mode), f_hat)
## LDA
roc_lda <- roc(as.factor(test_set$mode),p_hat_lda)
## QDA
roc_qda <- roc(as.factor(test_set$mode), p_hat_qda)
## Decision tree
roc_rpart <- roc(as.factor(test_set$mode), p_hat_rpart)
## Random forest
roc_rf <- roc(as.factor(test_set$mode), p_hat_rf)
# Graph with 7 ROC curves for each model
ggroc(list("Logistic regression" = roc_glm, "Naive Bayes" = roc_nb, "kNN" = roc_knn, "LDA" = roc_lda, "QDA" = roc_qda, "Decision tree" = roc_rpart, "Random forest" = roc_rf)) +
  theme(legend.title = element_blank()) +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color = "black", linetype = "dashed") +
  xlab("Sensitivity") +
  ylab("Specificity") 
```

###### AUC values

```{r}
auc(roc_glm)
auc(roc_nb)
auc(roc_knn)
auc(roc_lda)
auc(roc_qda)
auc(roc_rpart)
auc(roc_rf)
```

Among the 7 machine learning models, random forest has the highest overall accuracy (0.717) and AUC value (0.7496) ; and largest area under the ROC curve can be observed under the random forest model. Hence, the random forest model will be chosen as the best model to predict the mode of a song. 
