---
title: "Shiny_app_analysis"
author: "Tracy Cui"
date: "2021/12/10"
output: html_document
runtime: shiny
---

```{r}
library(shiny)
library(dplyr)
library(ggplot2)

ui <- fluidPage(
  # Application title
    titlePanel("Explore the Dataset"),
    
    h3("Distribution of Variables"),
    p("This serves as a tool for visual exploration of the variables in this dataset. We can also assess if any of the variables is biased, and refine our view of the distribution by adjusting number of bins."),
    tabsetPanel(
        tabPanel("Popularity",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of popularity. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hpop")
                     )
                 )),
        tabPanel("acousticness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of acousticness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num2",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("haco")
                     )
                 )),
        tabPanel("danceability",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of danceability. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num3",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hdan")
                     )
                 )),
        tabPanel("energy",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of energy. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num4",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hene")
                     )
                 )),
        tabPanel("instrumentalness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of instrumentalness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num5",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hins")
                     )
                 )),
        tabPanel("key",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of key.")
                     ),
                     mainPanel(
                         plotOutput("hkey")
                     )
                 )),
        tabPanel("liveness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of liveness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num6",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hliv")
                     )
                 )),
        tabPanel("loudness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of loudness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num7",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hlou")
                     )
                 )),
        tabPanel("speechiness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of speechiness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num8",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hspe")
                     )
                 )),
        tabPanel("valence",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of valence. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num10",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hval")
                     )
                 )),
        tabPanel("tempo",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of tempo. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num9",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("htem")
                     )
                 )),
        tabPanel("music_genre",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of music genre.")
                     ),
                     mainPanel(
                         plotOutput("hgen")
                     )
                 ))
    )
)

server <- function(input, output){
  data_clean <- readRDS("music_genre_clean.rds")
    data_clean$tempo <- as.integer(data_clean$tempo)
    top80 <- data_clean %>%
        filter(popularity >= 80)
    
    output$hpop <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(popularity), bins = input$num)
    })
    
    output$haco <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(acousticness), bins = input$num2)
    })
    
    output$hdan <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(danceability), bins = input$num3)
    })
    
    output$hene <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(energy), bins = input$num4)
    })
    
    output$hins <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(instrumentalness), bins = input$num5)
    })
    
    output$hkey <- renderPlot({
        ggplot(data_clean, aes(key)) +
            geom_bar()
    })
    
    output$hliv <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(liveness), bins = input$num6)
    })
    
    output$hlou <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(loudness), bins = input$num7)
    })
    
    output$hspe <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(speechiness), bins = input$num8)
    })
    
    output$htem <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(tempo), bins = input$num9)
    })
    output$hval <- renderPlot({
        ggplot(data_clean) +
            geom_histogram(aes(valence), bins = input$num10)
    })
    output$hgen <- renderPlot({
        ggplot(data_clean, aes(music_genre)) +
            geom_bar()
    })
}

shinyApp(ui = ui, server = server)


```

Popularity, which is the outcome variable for one of our tasks, takes integer value between 0 and 100 and represents the relative popularity of a song based on the number of weeks on a chart and survey score. It has two modes at around 48 and 55. Acousticness is between 0 and 1, where a song with acousticenss=1 uses non-electric instruments. The majority of the songs in this dataset are not acoustic. Danceability, (0, 1),  is a measure based on a song’s  tempo, rhythm, and beat that quantify how suitable it is for dancing. Higher danceability songs usually have stronger and more regular beats. Values of danceability in this dataset are close to normally distributed. Duration is the length of songs in milliseconds. Almost all songs in the dataset are less than 5 minutes. Energy quantifies the perceived intensity and activity of a song on a scale of 0.0 (low energy) to 1.0 (high energy). Distribution of Energy shows most of the songs are high energy. Instrumentalness depicts whether a track contains vocals, also on a scale of 0.0 (likely contains vocal content) to 1.0 (likely contains no vocal content). Liveness measures the presence of an audience in a song (hence whether a song is from a live performance), on a scale of 0.0 (no audience) to 1.0 (audible audience). Most liveness values are close to 0.0, which indicates the songs in our dataset are performed without an audience. Loudness is the average decibel level over a song. Speechiness quantifies the amount of spoken words in a song on a scale of 0.0 to 1.0. Distribution of speechiness is right skewed, meaning most songs have no or very little spoken words. Tempo is a common musical measure in beats per minute (bpm) of a song. Higher tempo corresponds to faster pace. Distribution of tempo values is very similar to that of "popularity".
Valence depicts the positivity, happiness and cheerfulness of a song on a 0-1 scale. Distribution of valence is right skewed, valence of most songs is between 0.2 and 0.7. Key is the major or minor scale around which a piece of music revolves. A song in a major key is based on a major scale. A song in a minor key is based on a minor scale. There are in total of 24 keys all of them are observed in our dataset. Mode of a song can be either major or minor, which describes the emotional character of a key. Usually, major tones are happier and lighter while minor tones express melancholy. But as music tends to be free of bondage, there are countless exceptions to this rule. 


```{r}
ui <- fluidPage(
  h3("Look at Top 20% Poplarity, Distribution of Variables"),
    p("One of the tasks of this project is to predict popularity. Here we visually assess the distribution of variables for the top 20% songs ranked by popularity." ),
    tabsetPanel(
        tabPanel("acousticness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of acousticness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num12",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("haco1")
                     )
                 )),
        tabPanel("danceability",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of danceability. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num13",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hdan1")
                     )
                 )),
        tabPanel("energy",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of energy. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num14",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hene1")
                     )
                 )),
        tabPanel("instrumentalness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of instrumentalness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num15",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hins1")
                     )
                 )),
        tabPanel("key",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of key.")
                     ),
                     mainPanel(
                         plotOutput("hkey1")
                     )
                 )),
        tabPanel("liveness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of liveness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num16",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hliv1")
                     )
                 )),
        tabPanel("loudness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of loudness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num17",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hlou1")
                     )
                 )),
        tabPanel("speechiness",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of speechiness. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num18",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hspe1")
                     )
                 )),
        tabPanel("valence",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of valence. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num20",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("hval1")
                     )
                 )),
        tabPanel("tempo",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of tempo. Use slider bar below to customize number of bins."),
                         
                         br(),
                         
                         sliderInput("num19",
                                     "Number of Bins:",
                                     value = 25,
                                     min = 1, max = 100)
                     ),
                     mainPanel(
                         plotOutput("htem1")
                     )
                 )),
        tabPanel("music_genre",
                 sidebarLayout(
                     sidebarPanel(
                         p("View distribution of music genre.")
                     ),
                     mainPanel(
                         plotOutput("hgen1")
                     )
                 ))
    )
)

server <- function(input, output) {
  data_clean <- readRDS("music_genre_clean.rds")
    data_clean$tempo <- as.integer(data_clean$tempo)
    top80 <- data_clean %>%
        filter(popularity >= 80)
    
  output$haco1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(acousticness), bins = input$num12)
    })
    
    output$hdan1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(danceability), bins = input$num13)
    })
    
    output$hene1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(energy), bins = input$num14)
    })
    
    output$hins1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(instrumentalness), bins = input$num15)
    })
    
    output$hkey1 <- renderPlot({
        ggplot(top80, aes(key)) +
            geom_bar()
    })
    
    output$hliv1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(liveness), bins = input$num16)
    })
    
    output$hlou1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(loudness), bins = input$num17)
    })
    
    output$hspe1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(speechiness), bins = input$num18)
    })
    
    output$htem1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(tempo), bins = input$num19)
    })
    output$hval1 <- renderPlot({
        ggplot(top80) +
            geom_histogram(aes(valence), bins = input$num20)
    })
    output$hgen1 <- renderPlot({
        ggplot(top80, aes(music_genre)) +
            geom_bar()
    })
}

shinyApp(ui = ui, server = server)

```

The variable distribution of the to 20% popularity songs shows popular songs have low levels of acousticness, instrumentalness, liveness, speechiness, and high levels of danceability, loudness. Hip-hop, Rap and Rock are the three most popular genres. C#, A3 and G# are the three most popular keys. 


```{r}
ui <- fluidPage(
    h3("Tempo, Dancibility, Liveness, Key and Popularity"),
    p("Based on our subject-mater knowledge, mainstream, popular music are usually \"upbeat\", 
      dancible and lively. Dancibility and liveliness might corresponse to certain trend in keys and tempos. 
      Here we plot dancibility and liveliness for songs across popularity, with each datapoint colored by the song's key and sized by its tempo"),
    
    sidebarLayout(
        sidebarPanel(
            # Add some text and a couple of hyper links before the slider for year
            p("Plot dancibility and liveliness for songs across popularity, with each datapoint colored by the song's key and sized by its tempo. Use slider bar to plot songs with a particular popularity."),
            
            # Add some space between the text above and animated
            # slider bar below
            br(),
            
            # Input: year slider with basic animation
            sliderInput("Popularity", "Popularity:",
                        min = 0, max = 99,
                        value = 0, 
                        step = 1,
                        sep = "",       # keep years in year format and not 1,960 format
                        ticks = FALSE,  # don't show tick marks on slider bar
                        animate = TRUE) # add play button to animate
        ),
        
        # Show a scatter plot for all countries
        mainPanel(
            plotOutput("scatterPlot")
        )
    )
    
)

# Define server logic required to draw a histogram
server <- function(input, output) {
  data_clean <- readRDS("music_genre_clean.rds")
    data_clean$tempo <- as.integer(data_clean$tempo)
    top80 <- data_clean %>%
        filter(popularity >= 80)
    output$scatterPlot = renderPlot({
        # Filter year to be the input year from the slider
        data_clean %>% filter(popularity %in% input$Popularity & !is.na(danceability)) %>%
            ggplot(aes(danceability, liveness, color = key, size = tempo)) +
            geom_point(alpha = 0.5) +
            xlab("Danceability") +
            ylab("Liveness") +
            
            # Set x and y axis limits to keep the same for each tempo
            scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
            scale_y_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
            
            # Make the legend titles nicer
            scale_color_discrete(name = "Key") +
            scale_size_continuous(name = "Tempo") +
            
            # Change the title of the plot for each tempo
            
            ggtitle(sprintf("Energy vs. Liveness for songs with popularity %d", input$Popularity)) 
    })
}

# Run the application 
shinyApp(ui = ui, server = server)

```

As popularity increments, we see a migration of the datapoint cluster towards high danceability and low liveness; there are less circles with smaller size on the plot, indicating faster tempo. In addition, we did not observe any particular prefernce for the musical key. Key selection is rather random and does not interfere with popularity.

